---
layout: post
title: "Visualizing the NBA's Shot Logs with R Part 1"
categories: [projects]
excerpt: Using R to scrape, clean, and visualize data from NBA.com's shot log API
tags: [NBA, data visualization, data science]
data: 2015-03-25
---

The NBA.com/stats site is arguably the best sports statistics interface across all sports and sports leagues. Thanks to a couple of great posts by Greg Reda and Daniel Forsyth ([here](gregreda.com/2015/02/15/web-scraping-finding-the-api/) and [here](danielforsyth.me/exploring_nba_data_in_python/)), I set out to build a wrapper in R. It's immediately clear that a set of unique player is needed and not easily accessible -- at least to me. In fact, there's probably an infinitely better way to get this data than the way I did, but alas...

My method: Brute force querying through all likely possible IDs. You can find my code and the complete table of over 5000 players in [this Github repo]().



```{r, message=FALSE, echo=FALSE}
library(knitr)
source("~/Projects/NBA-shots/shinyApp/shotLog.R")

```

You can also embed plots, for example:

```{r, echo=FALSE}
dat <- getShotLog("Stephen Curry", season = "2014-15")
heat.dat <- heat_data(dat)
result_map(heat.dat)
```

